MS Depression Paper, started 2022 08 24

### 2022-09-06 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220824_for_replication.Rmd
Add files via upload

---
title: "MS Depression Paper, started 2022 08 24"
author: "Erica Baller"
date: '2022-08-24'
output: html_document
---

@@

#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
#source("~/BBL/msdepression/scripts/wilcox_effsize.R")
```

Read in Data and do some preprocessing

```{r read_in_df_and_subset}
homedir <- "/Users/eballer/BBL/msdepression/"

@@

```
PROMIS DATA HANDLING to link a person to their most proximal PROMIS

@@

Demographics
```{r demographics}

@@

Define Depression Network by Volume of Overlap
```{r visualize_depression_network_volume}
###################################################
### Defining and Visualizing Depression Network ###
###################################################

@@

@@

dep_network_names_bottom_3_quartiles_no_cranial_nerves <- dep_network_names_bottom_3_quartiles_no_cranial_nerves %>%
   gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix
```

Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups

@@

dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000)

@@

@@

``` {r plots_for_dx_net_intx}
########## PLOTS #############

@@

@@

``` {r effect_size_regression}
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# lapply to get results of t-test in each fascicle
#ldepGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_cohen_d <- sapply(fascicle_names_w_vol_suffix, function(x) 
{
  text_for_eval <- paste0("cohensD(dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], dep_and_healthy_groups_for_ICD_analysis$", x, "[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_cohen_d) <- fascicle_names_w_vol_suffix

#convert to data frame
df_d_values <- as.data.frame(fascicle_dep_vs_healthy_cohen_d)

``` {r plot_scatter}
#######################
### plot differences ##
#######################

@@

df_fascicle_dep_mask_volume_overlap_d_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_d_values)
names(df_fascicle_dep_mask_volume_overlap_d_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_d")

# plot fascicle overlap x Eff size value
#statistic, p=001
dep_overlap_x_dep_vs_healthy_vol_d <- lm(dep_vs_healthy_d ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_d_value )

##### bootstrap

@@

#effect sizes
# 95% confidence intervals for the correlation
set.seed(1)
data_frame_dep_v_nondep_d_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_d_value, select = c("dep_vs_healthy_d", "volume_in_dep_mask"))

results_vol <- boot(data=data_frame_dep_v_nondep_d_x_vol_dep_mask, statistic=tval,
   R=10000, formula=dep_vs_healthy_d~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16

@@

scatter_d_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_d_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_d)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (D)") + 
  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", pval_vol)) + 
                 #round(summary(dep_overlap_x_dep_vs_healthy_vol_d )$coefficients[2,4],3))) + 
#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=12, color = "black"),
        axis.title = element_text(size=14, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank())
print(scatter_d_val_by_dep_net_overlap)

df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_d_value %>% dplyr::select(fascicle, dep_vs_healthy_d)
write.table(x = df_for_making_colors, file = paste0(homedir, "results/fascicle_and_effect_size_for_volxdx_analysis_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")

@@

@@

#### Save a data frame that contains the mean volume of disease within network, scaled by the size of the network, for visualization

#diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
#network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
#mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
 #                            mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))

#df_means_volume_scaled_by_network_size <-data.frame(diagnosis_vector, network_vector, mean_diagnosis_by_network_vector)
#names(df_means_volume_scaled_by_network_size) <- c("diagnosis", "network", "mean")
#write.table(df_means_volume_scaled_by_network_size, paste0(homedir, "/results/df_means_of_volumes_scaled_by_network_size_dx_net_for_color.csv"), row.names=F, col.names = T, quote = F, sep=",")

@@

#Diagnostic effects

#p=0.0328, dep = 0.4407, nondep = 0.387
msdisease_diagnostic_effects_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

#p=0.058, dep = 0.31, nondep= 0.28
msdisease_diagnostic_effects_wilcoxon_bottom_tertiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

#p=0.0325, dep = 0.4411, nondep = 0.386
msdisease_diagnostic_effects_wilcoxon_quintile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_quintile_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

#p=0.048, dep = 0.34, nondep = 0.30
msdisease_diagnostic_effects_wilcoxon_bottom_quintiles <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_quintiles_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])

>>>>

---
title: "MS Depression Paper, started 2022 08 24"
author: "Erica Baller"
date: '2022-08-24'
output:
  pdf_document: default
  html_document: default
---

## MS Depression

Welcome to the MS Depression Data! This project was designed to ask a relatively simple question - are patients with MS who are depressed more likely to have lesions in fascicles that connect areas previously associated with depression. The literature are sparse in this domain. It is very clear that depression occurs in up to 50% of patients with MS, and some studies have suggested that it relates to lesion burden. But to my knowledge there are no papers that evaluate how lesions disrupt white matter tracts that serve as the backbone for functional networks. Over the course of the next few chunks, I'll walk through our analysis. It combines imaging data from research grade scans acquired at 3T as part of a clinical protocol (T1 and T2 Flair), and clinical data from the electronic medical record.

The first chunk here is reading in appropriate data frames and preprocessing. A few big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller.

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

### Inclusion/Exclusion

#### Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)

  1) Scanned under MS protocol (42K)

  2) MS Diagnosis (17K)

  3) Seen by MS doctor (16K)

#### Inclusion criteria for Depressed Group

  1) Have an ICD10 F3* code (in our sample, people have F32 (depressive episode), F33 (major depressive disorder), and F34 (Persistent mood [affective] disorders)) 
        - there are NO participants with bipolar disorder, manic episode, or unspecified mood/affective disorder)
        - while participants may have been diagnosed with F06.3 (Depressive disorder due to known physiological condition, with depressive features), they must ALSO have had an F3(2-4) condition

  2) Screened positive for depression on PHQ2 or PHQ9 at any point

  3) Prescribed antidepressant medications per NAMI website
        - https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

#### Inclusion criteria for Nondepressed Group

  1) No F* diagnoses

  2) At least 1 PHQ2 or PHQ9 with a documented 0

  3) Free of psychiatric medications per NAMI website

#### EMR Exclusions

  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

#### Imaging exclusion

  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group. We are specifically using their FIRST scan.

### Group validation and depression network definition

#### Functional Impairment Scores

  To validate our grouping, we wanted to compare symptom burden between patients with depression and nondepressed comparators. One challenge of EMR research is that we are limited to what has been coded in particular fields, and could not readily assess our patients, or access their notes. As PHQ2 and 9 were used in dignostic categorization, they, by default differ between groups and are less informative. MS impairment is often assessed and recorded via EDSS, but this is not currently coded in the EMR. However, the Penn EMR has started recording Patient-Reported Outcomes Measurement Information System (PROMIS). Per the NIH, "PROMIS addressed a need in the clinical research community for a rigorously tested patient reported outcome (PRO) measurement tool that uses recent advances in information technology, psychometrics, and qualitative, cognitive, and health survey research to measure PROs such as pain, fatigue, physical functioning, emotional distress, and social role participation that have a major impact on quality-of-life across a variety of chronic diseases." PROMIS scores, for reference, are scored where higher numbers mean better functioning, and lower numbers mean worse functioning.

  A small subset of our participents (36 in non-depressed group and 49 in depressed group) had PROMIS scores. For participants who had multiple PROMIS scores, the score that was analyzed was the closest in time proximity to the day of the scan. PROMIS score cleaning and linking of score to scan is performed in a separate markdown **(promis_score_cleaning_2010-2022.Rmd). **

  Based on previous MS literature, we hypothesized that our patients with depression would overall show more functional impairment than patients without depression.

#### Depression Network Construction

  A note on the "depression network." This network was made by Shan Siddiqi in Michael Fox's group by doing a conglomeration of TMS datasets, stroke datasets, Nature Human Behavior 2021 (https://www.nature.com/articles/s41562-021-01161-1). It is an unthresholded mask (he initially sent one where the unit is r, but then sent us one where unit is T). 
  Per Siddiqi's article: The location of each lesion or brain stimulation site (Fig. 2a–c, top panels) was mapped to an underlying brain circuit using a large normative connectome database (n=1000) and previously validated methods (Fig. 2a–c, bottom panels)5. The normative connectome was used to estimate connectivity of each lesion or stimulation site to every voxel in the brain. At each voxel, a Pearson r-value was computed for the correlation between depression score and lesion or stimulation site connectivity to that voxel (Fig. 2a–c, right panels), yielding a population-derived “circuit map” for each of the 14 datasets (Supplementary Figure 1). We generated a combined depression circuit map by taking the mean of all 14 circuit maps, weighted by the sample size of each dataset (Fig. 5a). Of note, lesions are expected to positively correlate with symptoms, whereas the inverse of tms sites and was (so connectivity in these areas was inverted)

  To use it, I thresholded the clusters (3.09), binarized it and then used it as an ROI and calculated, per fascicle, the volume occupied by the fascicle that intersected with the depression mask. Most fascicles were either entirely overlapping or entirely outside the depression mask, though the range of overlap was anywhere from 1-30% (volume of fascicle overlap/total volume of fascicle). This required us to make some choices. We considered an option where the "depression mask" was anywhere that ever overlapped, and non-depression were areas where there was no overlap. We also looked at quartiles, and settled on the top 25% (top quartile), i.e. the top 25% of fascicles with the highest volume of network overlap were considered in the depression network. Everything outside of that was considered "non_depression" network, or nondep_net. Cranial nerves are too small to reliably be assess with our method, and so they were removed from the analysis.In total, 77 fascicles were evaluated.

### Analyses
  Demographic differences between groups were assessed with T-tests and chi-square for categorical variables w/the command CreateTableOne.

#### *Disease burden measures per individual (done in the shell):*

    For each individual, the number of 1s in the MIMoSA binary mask were summed to get a measure of overall lesion burden (to be distinguished from fascicle burden).

    To assess fascicle burden for each individual, streamline filtering was performed.

    Streamline filtering is an interative process performed in DSI studio. For each individual, the MIMoSA binary map was considered a region of interest. For each of the 77 fascicles, streamlines that ran through the lesion were "filtered" or kept, whereas the fascicles that avoided the MIMoSA mask were eliminated. Streamlines that passed through the MIMoSA were then saved binary .nii files, where 1 indicated that disease was present in that voxel, and 0 indicated either 1) that fascicle did not cross through that voxel or there was no disease in it.

    I was then able to calculate the "volume" of the disease in a fascicle (i.e. volume of the streamlines that were affected) by summing the # of 1s in the map. At the end, each individual had 77 single values that represented the volume of affected streamlines within each fascicle.

    Full fascicle volumes were also calculated and saved as .niis.

####  *Disease burden summary measures (done in this R script):*

    Having computed disease measures at the individual fascicle, I wanted to assess network effects for all future analyses. To do this, I calculated three summary measures per individual.

    1) Total disease : Sum of all 77 volume measures of disease, divided by the total volume of "healthy" fascicles (i.e. taking a sum of all of the full fascicle volumes). This yields a proportion of the overall burden of disease in the brain

    2) Depression network: Sum of all 19 volume measures of disease in the fascicles within the depression network, divided by the sum of total full fascicle volumes in the depression network.

    3) Nondepression network: Sum of all 58 volume measures of disease in the fascicles outside the depression network, divided by the sum of total full fascicle volumes in the nondepression network.

####  *Network effects:*

   Given that depression is so common in multiple sclerosis, we first wanted to check whether disease was diffuse across the brain, or whether there was network specificity (i.e. is MS random, or do MS lesions happen to strike fascicles that support the depression network, irrespective of diagnosis). To assess this, I took the following steps.

    1) Tests for normality (Shapiro–Wilk test) of proportion of disease in depression network and nondepression network. Both not normally distributed

    2) Used wilcoxon tests to compare disease (paired = T) between depression and nondepression network. As each individual contributed one measure for each network (i.e. 2 measures per person), paired was used

####  *Diagnostic effects:*

  Per our inclusion/exclusion criteria, we were able to identify a subset of patients that had depression, as well as a subset who did not meet criteria per EMR review. We next wanted to assess whether patients with depression had more overall disease than patients without depression.

    First, we assessed whether lesion burden (i.e. the volume occupied by white matter disease), irrespective of which fascicles were impacted, differed between patients and controls.

    1) Tests for normality (Shapiro–Wilk test), volume of lesions not normally distributed.

    2) Used wilcoxon tests to compare lesion burden between depressed and non-depressed patients. As each person only contributed one measure, paired = F.

  Next, we assessed whether fascicle burden (i.e. total volume of disease divided by total volume of all fascicles) differed between diagnostic groups.

    1) Tests for normality (Shapiro–Wilk test), fascicle burden not normally distributed.

    2) Used wilcoxon tests to compare fascicle burden between depressed and non-depressed patients. As each person only contributed one measure, paired = F.

  We next aimed to see whether there was an interaction of Diagnosis and Network. Specifically, we hypothesized that patients with depression would have more disease specifically in the depression network.

     1) Model: diagnosis_by_network_interaction_lmer <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions)

  Given the somewhat arbitrary definition of depression network (25%/75%), we next assessed whether the relationship between diagnosis and network was continuous.

    1) For each fascicle, two values were computed

      a) Effect size (r) from the wilcoxon statistic comparing volume of disease in the fascicle between depressed vs nondepressed

      b) The volume of the overlap of that fascicle with the depression network

    2) A linear model relating the overlap of volume of the fascicle w/the depression network to the effect size from the depressed v nondepressed wilcoxon analysis.

    3) This relationship was permuted 10,000 w/boot (w/replacement)

    4) Plots were generated

####  *Sensitivity Analyses*

  Though the effect size by overlap analysis was reassuring, we wanted to further guarantee that our findings were not driven by arbitrary distinctions between the depression and non-depression networks. To assess this, we repeated the network analysis using tertiles and quintiles. LMER for assessing interactions was used.

    1) Tertiles:

      a) Depression network: Top Tertile
      b) Nondepression network: Bottom 2 Tertiles

    2) Quintiles:

      a) Depression network: Top Quintile
      b) Nondepression network: Bottom 4 Quintiles

## Results:

#### Demographics

232 with depression (depGroupVar == 2), 148 healthy (depGroupVar == 1)

                         Healthy          Depressed             p      test
  n                                148            232                     
  Race (%)         Caucasian       108 ( 73.0)    173 ( 74.6)   0.821     
                   Non-caucasian    40 ( 27.0)     59 ( 25.4)             
  Sex (%)          Female          117 ( 79.1)    199 ( 85.8)   0.117     
                   Male             31 ( 20.9)     33 ( 14.2)             
  Age (mean (SD))                46.57 (12.66)  48.75 (11.75)   0.089     
  Depression (%)   1               148 (100.0)      0 (  0.0)  <0.001     
                   2                 0 (  0.0)    232 (100.0)             
  PHQ2 (mean (SD))                0.00 (0.00)    0.58 (1.53)   <0.001     
  PHQ9 (mean (SD))                0.00 (NA)     10.43 (9.02)       NA

#### PROMIS Scores
                                               level   1              2              p      test
  n                                                       36             49                     
  Quality_of_Life (mean (SD))                           3.59 (1.02)    2.95 (0.99)    0.010     
  Physical_Health (mean (SD))                           2.86 (0.83)    2.41 (0.76)    0.019     
  Mental_Health_and_Mood (mean (SD))                    3.97 (0.82)    3.39 (0.99)    0.011     
  Social_Activities_Satisfaction (mean (SD))            3.83 (1.00)    2.84 (1.36)    0.001     
  Carrying_Out_Social_Activities (mean (SD))            3.71 (1.12)    2.93 (1.21)    0.007     
  Carrying_Out_Physical_Activities (mean (SD))          3.68 (1.06)    3.14 (1.00)    0.032     
  Emotional_Problems (mean (SD))                        3.67 (1.00)    3.00 (1.08)    0.011     
  **Fatique_Average (mean (SD))                           3.48 (1.25)    3.09 (1.14)    0.181  *   
  PostOp.PROMIS.Physical.Score (mean (SD))             13.15 (3.03)   11.67 (2.55)    0.032     
  PostOp.PROMIS.Mental.Score (mean (SD))               15.00 (2.92)   12.31 (3.31)    0.001

  This is actually great news - if depressed/nondepressed individuals differed on fatigue, you could make the case that of COURSE they'd feel worse as a consequence of their fatigue. However, that they weren't different here lends some specificity to the measures.

#### Depression Network Construction

dep_network_names_top_quartile_no_cranial_nerves
 [1] CC      AF_L    SLF2_R  SLF3_R  FAT_L   TR_S_R  SLF2_L  AF_R    TR_S_L  FAT_R   CS_S_R  CPT_F_R CPT_P_L ILF_R   CPT_P_R CS_S_L  PAT_R   MdLF_L  CBT_R

dep_network_names_bottom_3_quartiles_no_cranial_nerves
  [1] C_FPH_L C_FPH_R C_FP_L  C_FP_R  C_PH_L  C_PHP_L C_PHP_R C_PH_R  C_R_L   C_R_R   EMC_L   EMC_R   IFOF_L  IFOF_R  ILF_L   MdLF_R  PAT_L   SLF1_L  SLF1_R 
[20] SLF3_L  UF_L    UF_R    VOF_L   VOF_R   CB_L    CB_R    ICP_L   ICP_R   MCP     SCP     V       AR_L    AR_R    CBT_L   CPT_F_L CPT_O_L CPT_O_R CS_A_L 
[39] CS_A_R  CS_P_L  CS_P_R  CST_L   CST_R   DRTT_L  DRTT_R  F_L     F_R     ML_L    ML_R    OR_L    OR_R    RST_L   RST_R   TR_A_L  TR_A_R  TR_P_L  TR_P_R 
[58] AC

#### Network Effects

    - Disease within versus outside depression network: p = 2.2*10^-16, dep net prop = 0.400, non dep = 0.339, eff size 0.73

#### Diagnostic Effect

    - Volume of MIMoSA lesion burden: p=0.036; depressed = 15,871; nondepressed = 12,886, eff size 0.107

    - Proportion Volume of whole brain fascicle disease burden: W=14995, p=0.04, dep = 0.39, nondep = 0.30

#### Network x Diagnosis Interaction

    - Main effect of Diagnosis: T = 2.29, p = 0.02 (Depressed patient disease > Nondepressed patient disease)

    - Main effect of Network: T = 16.58, p <2.2 * 10^-16 (Depression Network more disease than nondepression network)

    - Interaction= T = 2.2, p = 0.03; finding driven by worse disease in depressed patients, within depression network

#### Effect Size x Fascicle overlap with depression network scatter

    - p before bootstrap: p=1.8*10^-6; effect increases as overlap with depression network increases

    - p after bootstrap: p = < 1*10^-16

#### Sensitivity Analyses

      1) Tertiles

        - Main effect of network (Depression > Nondepresison network): T=23.2, p = 2*10^-16

        - Main effect of diagnosis (Depressed > Nondepressed people):  T = 2.37, p = 0.018

        - Interaction of network and diagnosis:  T = 2.17, p = 0.031

      2) Quintiles

        - Main effect of network (Depression > Nondepresison network): T=22.56, p = 2*10^-16;

        - Main effect of diagnosis (Depressed > Nondepressed people): T = 2.285, p = 0.019

        - Interaction of network and diagnosis, T = 2.34, p = 0.020

@@

homedir <- "/Users/eballer/BBL/msdepression/"

#functions
source(paste0(homedir,"scripts/msdepression_functions.R"))

### Read in Data and do some preprocessing

```{r read_in_df_and_subset}

@@

```
### PROMIS DATA HANDLING to link a person to their most proximal PROMIS

@@

### Demographics
```{r demographics}

@@

### Define Depression Network by Volume of Overlap
```{r visualize_depression_network_volume}
###################################################
### Defining and Visualizing Depression Network ###
###################################################

@@

write.table(dep_network_names_top_quartile_no_cranial_nerves, paste0(homedir, "/results/dep_network_names_top_quartile_no_cranial_nerves_by_vol.txt"), row.names = F, col.names = F, quote = F)

df_for_making_binary_color_maps <- fascicle_volumes_dep_network_no_cranial_nerves %>% dplyr::select(fascicle, non_zero_voxels_in_dep_map)

write.table(df_for_making_binary_color_maps, paste0(homedir, "/results/volume_of_overlap_w_dep_net_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")

@@

dep_network_names_bottom_3_quartiles_no_cranial_nerves <- dep_network_names_bottom_3_quartiles_no_cranial_nerves %>%
   gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE) #n=68, w/volume suffix

### Compare overall volume of the lesions, irrespective of fascicle involvement, between diagnostic groups

@@

dep_vs_healthy_wilcoxon_lesion_volume_eff_size <- rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = volume_of_mimosa_lesions ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = "perc", nboot = 1000) #eff size 0.107 (small)

@@

### Compare between networks, diagnostic groups, and interaction

@@

```
### Plot interaction
``` {r plots_for_dx_net_intx}
########## PLOTS #############

@@

```
### Effect size regression (effect size x overlap with dep net)
``` {r effect_size_regression}
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles_prop_and_vol %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_vol_lost = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

# 
#depGroup 2 is depressed group, Dep group 1 is healthy group. Positive Ts are places where depressed people have more disease than controls
fascicle_dep_vs_healthy_effsize <- sapply(fascicle_names_w_vol_suffix, function(fascicle) 
{

  text_for_eval <- paste0("rstatix::wilcox_effsize(dep_and_healthy_groups_for_ICD_analysis, formula = ", fascicle, " ~ depGroupVar, paired = F, conf.level = 0.95, ci.type = \"perc\", nboot = 1000)$effsize")
  eval(parse(text=text_for_eval))
})
names(fascicle_dep_vs_healthy_effsize) <- fascicle_names_w_vol_suffix

#convert to data frame
df_effsize_values <- as.data.frame(fascicle_dep_vs_healthy_effsize)

write.table(x = df_effsize_values, file = paste0(homedir, "results/fascicle_and_wilcox_effect_size_for_volxdx_analysis_n77.csv"), row.names = T, col.names = F, quote = F, sep = ",")

@@

### Plot effect size x depression net overlap
``` {r plot_scatter}
#######################
### plot differences ##
#######################

@@

df_fascicle_dep_mask_volume_overlap_effsize_value <- data.frame(fascicle_volumes_dep_network_no_cranial_nerves$fascicle, fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, df_effsize_values)
names(df_fascicle_dep_mask_volume_overlap_effsize_value) <- c("fascicle", "volume_in_dep_mask", "dep_vs_healthy_effsize")

# plot fascicle overlap x Eff size value
#statistic, p=1.8*10^-6
dep_overlap_x_dep_vs_healthy_vol_effsize <- lm(dep_vs_healthy_effsize ~ volume_in_dep_mask, data=df_fascicle_dep_mask_volume_overlap_effsize_value )

@@

#effect sizes
# 95% confidence intervals for the correlation
set.seed(1)
data_frame_dep_v_nondep_effsize_x_vol_dep_mask <- subset(df_fascicle_dep_mask_volume_overlap_effsize_value, select = c("dep_vs_healthy_effsize", "volume_in_dep_mask"))

results_vol <- boot(data=data_frame_dep_v_nondep_effsize_x_vol_dep_mask, statistic=tval,
   R=10000, formula=dep_vs_healthy_effsize~volume_in_dep_mask)
print(results_vol)
plot(results_vol)
boot_results_vol.ci <- boot.ci(results_vol, conf=0.95, type="perc", index=1)
print(boot_results_vol.ci)
pval_vol<-boot.pval(results_vol, type = "perc", theta_null = 0, pval_precision = NULL)
print(pval_vol) #p=1*10^-16; CI (1.774-5.144)

@@

scatter_effsize_val_by_dep_net_overlap <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize)) +
  geom_point(size=1) + 
  geom_smooth(method=lm)+
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (r)") + 
  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network, p = ", pval_vol)) + 
                 #round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize )$coefficients[2,4],3))) + 
#  ggtitle(paste0("Effect of Diagnosis by Overlap with Depression Network")) + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=12, color = "black"),
        axis.title = element_text(size=14, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank())
print(scatter_effsize_val_by_dep_net_overlap)

df_for_making_colors <- df_fascicle_dep_mask_volume_overlap_effsize_value %>% dplyr::select(fascicle, dep_vs_healthy_effsize)
write.table(x = df_for_making_colors, file = paste0(homedir, "results/fascicle_and_effect_size_for_volxdx_analysis_n77.csv"), row.names = F, col.names = F, quote = F, sep = ",")

@@

### Sensitivity Analyses

@@

@@

#################
## Interaction ##
#################

#### tertile
df_empi_dx_network_proportions_tertiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_tertile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_tertiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#diagnosis: Dep > nondep T = 2.37, p = 0.018; Network dep>nondep = T=23.2, p = 2*10^-16; intx = T = 2.17, p = 0.031
diagnosis_by_network_interaction_lmer_tertile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_tertiles)
summary(diagnosis_by_network_interaction_lmer_tertile)

#### quintile
df_empi_dx_network_proportions_quintiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_quintile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_quintiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

#diagnosis: Dep > nondep T = 2.285, p = 0.019; Network dep>nondep = T=22.56, p = 2*10^-16; intx = T = 2.34, p = 0.020
diagnosis_by_network_interaction_lmer_quintile <- lmer(proportion_affected ~ Diagnosis*Network + (1|subject), data = df_empi_dx_network_proportions_quintiles)
summary(diagnosis_by_network_interaction_lmer_quintile)
